---
title: "Reproducible Research: Assessment 1"
author: Dapeng
output: 
  html_document:
    keep_md: true
---
## First, identify the library need for the project
```{r}
library("plyr")
library("knitr")
library("reshape2")
library("ggplot2")
```

## Loading and preprocessing the data
The data is a .csv file and saved in "activity.csv".
```{r}

File_Path <- "./activity.csv"
data = read.table(file = File_Path,header=T, quote="\"", sep=",")
data$date <- as.Date(data$date)
head(data)
```

## Question 1: What is mean total number of steps taken per day?
We will ignor the NA values in the steps

1. We will calculate the number of steps taken per day
2. Make a Histogram plot on the steps per day for the past 2 months
3. Calculate the mean value

```{r}
data_NO_na <- melt(data = data,id.vars="date", measure.vars="steps", na.rm= TRUE)
steps_date <- dcast(data_NO_na, date ~ variable, sum)
plot(steps_date$date, steps_date$steps, type="h", main="Histogram of Daily Steps",xlab="Date", ylab="Steps per Day", col="black", lwd=5)
abline(h=mean(steps_date$steps), col="red", lwd=2)
```

The Mean steps per day is:
```{r echo =FALSE}
mean(steps_date$steps)
```
The Median steps per day is:
```{r echo=FALSE}
median(steps_date$steps)
```

## Question 2: What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
data_NO_na <- melt(data = data,id.vars="interval", measure.vars="steps", na.rm= TRUE)
interval_date <- dcast(data_NO_na, interval ~ variable, mean)
head(interval_date)
plot(interval_date$interval, interval_date$steps, type="l", main="time series Steps",xlab="interval", ylab="Average Number of Steps", col="black", lwd=2)
abline(h=mean(interval_date$steps), col="red", lwd=2)
```

Report the interval for max steps
```{r}
h=interval_date$interval[which(interval_date$steps == max(interval_date$steps))]
max<-max(interval_date$steps)
paste("Interval with max steps is", h)
paste("The max steps for interval is", max)
```

https://rpubs.com/Liang/Reproducible_Research_porject_1
https://rstudio-pubs-static.s3.amazonaws.com/59476_8947b783031e47b38c4fa74ab11345ac.html


##Question 3: Imputing missing values
First, take a look on how many days with missing values.
```{r}
sum(is.na(data$steps))

```

Secone, i will replace the NA with the mean steps for the 5 min interval. The new data is saved in data_ReplaceNA
```{r}
meansteps<-mean(data$steps,na.rm = T)
data_ReplaceNA<-data
data_ReplaceNA$steps[is.na(data_ReplaceNA$steps)]=meansteps
head(data_ReplaceNA)
```

Calculate total number of steps taken each day
```{r}
df_steps_by_day<-aggregate(steps~date,data_ReplaceNA,sum)
head(df_steps_by_day)
hist(df_steps_by_day$steps, main="Histogram of total number of steps per day", xlab="Total number of steps in a day", breaks=15)
```

##Question 4: Are there differences in activity patterns between weekdays and weekends?
1. Add a new factor variable "type_of_day" in the data frame
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data

```{r}
data_ReplaceNA$type_of_day<-weekdays(as.Date(data_ReplaceNA$date))
head(data_ReplaceNA)
data_ReplaceNA$type_of_day[data_ReplaceNA$type_of_day %in% c("Saturday","Sunday")]<-"Weekend"
data_ReplaceNA$type_of_day[data_ReplaceNA$type_of_day !="Weekend"]<-"Weekday"
data_ReplaceNA$type_of_day=as.factor(data_ReplaceNA$type_of_day)
data_noNA_mean<-aggregate(steps~interval+type_of_day,data_ReplaceNA,mean)
ggplot(data_noNA_mean,aes(x=interval, y=steps))+geom_line()+facet_wrap(~type_of_day)
```

The step in the weekend is more evenly distributed in the intervals for the weekend, while more steps are walked in the morning.
